---
title: "Homework 4: Logistic regression and judges"
author: "Megan Blanchard"
date: "Due: Sunday, April 19, 2015"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
library(foreign)
library(dplyr)
library(ggplot2)

data <- read.dta("http://data.princeton.edu/wws509/datasets/greene.dta")

   
```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
data %>%
  select(decision) %>%
  mutate(decision.dummy = ifelse(decision == "no", 0,
                          ifelse(decision == "yes", 1, NA))) %>%
  summarise(prop.granted = mean(decision.dummy))
```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
data.prop.granted <- data %>%
  select(decision, judge) %>%
  mutate(decision.dummy = ifelse(decision == "no", 0,
                          ifelse(decision == "yes", 1, NA))) %>%
  group_by(judge) %>%
  summarise(prop.granted = mean(decision.dummy)) %>%
  arrange(prop.granted)

data.prop.granted 
```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
ggplot(data= data.prop.granted, aes(judge, prop.granted)) + geom_point()
```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whetheran appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
data <- data %>%
  mutate(decision.dummy = ifelse(decision == "no", 0,
                          ifelse(decision == "yes", 1, NA)),
         judge= as.factor(judge))
data$judge <- relevel(data$judge, ref="Iacobucci")
reg <- glm(decision.dummy ~ judge, family="binomial", data=data)


```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
pp <- coef(reg)
plogis(pp[1] + pp["judgeMarceau"])
#the predicted probability that Judge Marceau will grant an appeal is 60% 
   
```

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
data$city <- relevel(as.factor(data$city), ref="other")
data$language <- relevel(as.factor(data$language),ref="French")
reg2 <- glm(decision.dummy ~ judge + merit + city + language, family=binomial, data=data)


```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
pp2 <- coef(reg2)
plogis(pp2[1] + pp2["judgeMarceau"] + pp2["merityes"] +pp2["languageEnglish"] +pp2["cityToronto"])
#the predicted probability that Judge Marceau will grant an appeal for a case that was ruled to have merit, that took place in Toronto, and that was in English  is 90% 
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


