---
title: "Homework 4: Logistic regression and judges"
author: "Grace Tien"
date: "Due: Sunday, April 19, 2015"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
library(dplyr)
library(ggplot2)
library(broom)

GS_data <- read.table("data/Greene.txt")
names(GS_data)
```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
yes <- GS_data %>%
  filter(decision == "yes") %>%
  summarise(yes = n())
  
prop_yes <- yes/length(GS_data$decision)
prop_yes
```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
cases <- GS_data %>%
  group_by(judge) %>%
  summarise(cases = n())
cases

p <- ggplot(data=cases, aes(x=judge, y=cases)) + geom_bar(stat="identity") + labs(x="Judge", y="Cases") + ggtitle("Cases per Judge")
p
```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
rate_grants <- GS_data %>%
  group_by(judge, decision) %>%
  filter(decision=="yes") %>%
  summarise(yes = n())

cases$yes <- rate_grants$yes

cases1 <- cases %>%
mutate(rate_grants=yes/cases)

p1 <- ggplot(data=cases1, aes(x=judge, y=rate_grants), color="green") + geom_bar(stat="identity") + labs(x="Judge", y="Rate") + ggtitle("Rate of Granted Appeals")
p1
```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whetheran appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
GS_data$judge <- relevel(as.factor(GS_data$judge), ref = "Iacobucci")
fit1 <- glm(decision ~ judge, data = GS_data, family=binomial)
fit1 <- tidy(fit1)
fit1
```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
plogis(fit1$estimate[1] + fit1$estimate[7])
```

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
#relevel
GS_data$location <- relevel(as.factor(GS_data$location), ref="other")
GS_data$language <- relevel(as.factor(GS_data$language), ref="French")

#fit a model
fit2 <- glm(decision ~ judge + rater + location + language, data = GS_data, family=binomial)
fit2 <- tidy(fit2)
fit2
```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
plogis(fit2$estimate[1] + fit2$estimate[7] + fit2$estimate[11] + fit2$estimate[14] + fit2$estimate[13])
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


