---
title: "Homework 4: Logistic regression and judges"
author: "Kyonne"
date: "Due: Sunday, April 19, 2015"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(broom))

dat <- read.table("data/Greene.txt", header=TRUE)
#View(dat)
```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
dat.decide.all <- dat %>%
  group_by(judge) %>%
  summarise(all = n())

dat.appeal <- dat %>%
  group_by(judge) %>%
  filter(decision == "yes") %>%
  summarise(decision = n())

appeal.table <- full_join(dat.decide.all, dat.appeal, by = "judge") %>%
  mutate(prop.appeal = decision / all)

```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
#I only count 10 different judges in the data!
dat %>%
  group_by(judge) %>%
  summarise(n())

#Graph:
graph1 <- ggplot(data = dat.decide.all, 
                 aes(x = judge, y = all, fill = judge))
graph1 +  geom_bar(stat = "identity", 
           colour = "black", 
           position="dodge",
           show_guide = FALSE) +
  xlab("Judge Name") +
  ylab("Number of Cases") +
  ggtitle("Number of Cases for Each Judge") +
  theme_bw() 
```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
graph2 <- ggplot(data = appeal.table, 
                 aes(x = judge, y = prop.appeal, fill = judge))
graph2 +  geom_bar(stat = "identity", 
           colour = "black", 
           width = 0.7, 
           position="dodge",
           show_guide = FALSE) +
  xlab("Judge Name") +
  ylab("Proportion of Cases Granted an Appeal") +
  ggtitle("Proportion of Cases\nGranted an Appeal for each Judge") +
  theme_bw() 
```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whetheran appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
#Change reference level:
levels(dat$judge)
dat2 <- within(dat, judge <- relevel(judge, ref = "Iacobucci"))
##check that levels have been adjusted:
levels(dat2$judge)

#Logistic regression:
m1 <- glm(decision ~ judge, data = dat2, family = "binomial")
m1 <- tidy(m1)
m1
```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
(exp(m1$estimate[1] + m1$estimate[7])) / (1 + exp(m1$estimate[1] + m1$estimate[7]))

```

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
#Relevel factors:
dat3 <- within(dat2, location <- relevel(location, ref = "other"))
dat4 <- within(dat3, language <- relevel(language, ref = "French"))

#Logistic regression:
m2 <- glm(decision ~ judge + rater + location + language, 
          data = dat4, family = "binomial")

m2 <- tidy(m2)
m2
```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
plogis(m2$estimate[1] + m2$estimate[7] + m2$estimate[11] + m2$estimate[13] + m2$estimate[14])
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


