---
title: "Homework 4: Logistic regression and judges"
author: "Aaron Kurosu"
date: "Due: Wednesday, April 22, 2015 (extension)"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
library(dplyr)
library(ggplot2)
library(broom)



dat <- read.table("data/Greene.txt", header=TRUE)

```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
# languages less likely to have an appeal granted?
dat %>%
  select(language, decision) %>%
  group_by(language) %>%
  summarise(yes = mean(decision=="yes")) %>%
  ggplot(aes(x=language, y=yes)) +
    geom_bar(stat="identity") +
    ylab("percent appealed granted")

# cases vs. percent appeal granted per nation:
dat %>%
  select(nation, decision) %>%
  ggplot(aes(x=nation, fill=decision)) +
    geom_bar() +
    theme(axis.text.x=element_text(angle=90, hjust=1))

```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
# are the judges fair 50/50 in granting appeal?
dat %>%
  select(judge, decision) %>%
  group_by(judge) %>%
  summarise(cases = sum(decision)) %>%
  ggplot(aes(x=judge, y=cases, label=cases)) +
    geom_bar(stat="identity") +
    geom_text(vjust=-1) +
    expand_limits(y=c(0, 100)) +
    ylab("number of cases")
```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
# are the judges fair 50/50 in granting appeal?
dat %>%
  select(judge, decision) %>%
  group_by(judge) %>%
  summarise(yes = mean(decision=="yes")) %>%
  ggplot(aes(x=judge, y=yes)) +
    geom_bar(stat="identity") +
    expand_limits(y=c(0, 1)) +
    ylab("percent of appeals granted")
```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whetheran appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
# set judge Iacobucci as reference category :
dat$judge <- relevel(dat$judge, "Iacobucci")

glm(decision ~ judge, data=dat, family="binomial") %>%
  summary()
```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
print("When compared to judge Iacobucci, the probability judge Marceau will grant an appeal is:")

glm(decision ~ judge, data=dat, family="binomial") %>%
  tidy() %>%
  filter(term == "(Intercept)" | term == "judgeMarceau") %>%
  select(term, estimate) %>%
  summarise(percent = round((estimate[2]+estimate[1])*100, digits=2))
```

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
# set other as reference category:
dat$location <- relevel(dat$location, "other")

# Dummy variable for language set as English because French is treated as ommitted category, i.e. when english isn't true:
dat <- dat %>%
  mutate(english = as.integer(language=="English"))

glm(decision ~ judge + rater + location + english, 
    data=dat, family="binomial") %>%
  summary()
```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
print("When compared to judge Iacobucci, the probability judge Marceau will grant an appeal if the case has merit, took place in Toronto, and was in English is:")

glm(decision ~ judge + rater + location + english, 
    data=dat, family="binomial") %>%
  tidy() %>%
  filter(term == "(Intercept)" | term == "judgeMarceau" | 
           term == "locationToronto" | term =="english") %>%
  select(term, estimate) %>%
  summarise(percent = round((estimate[2]+estimate[1]+
                               estimate[3]+estimate[4])*100, digits=2))
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


