---
title: "Homework 4: Logistic regression and judges"
author: "Jessamin Birdsall"
date: "Due: Sunday, April 19, 2015"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
library(dplyr)
library(ggplot2)
library(broom)
data <- read.table("data/Greene.txt")
```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
head(data)

yes <- data %>%
  filter(decision == "yes") %>%
  summarise(yes <- n())
  
prop_yes <- yes/length(data$decision)
prop_yes
```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
cases <- data %>%
  group_by(judge) %>%
  summarise(cases = n())
p <- ggplot(data = cases, aes( x = judge, y = cases), color = "blue") + geom_bar(stat = "identity") +
  labs(x = "Judge", y = "Number of Cases") +
  ggtitle("Number of Cases Per Judge")
p

```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
rate_yes <- data %>%
  group_by(judge, decision) %>%
  filter(decision == "yes") %>%
  summarise(yes = n()) 

cases$yes <- rate_yes$yes

cases <- cases %>%
  mutate(rate_yes = (yes/cases))

p1 <- ggplot(data = cases, aes(x = judge, y = rate_yes)) + 
  geom_bar(stat = "identity") +
  labs(x = "Judge", y = "Rate of Granting Appeals") +
  ggtitle("Rate of Granting Appeals per Judge")
p1

```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whether an appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
data$judge <- relevel(as.factor(data$judge), ref="Iacobucci")
fit <- glm(decision ~ judge, family = "binomial", data = data)
tidy(fit)
```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
b <- coef(fit)
b.marceau <- b["judgeMarceau"]
exp(b.marceau)
plogis(b[1] + b.marceau)
```
The predicted probability that Judge Marceau will grant an appeal is 60%.

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
data$location <- relevel(as.factor(data$location), ref = "other")
data$language <- relevel(as.factor(data$language), ref = "French")
fit2 <- glm(decision ~ judge + rater + location + language, data = data, family = binomial)
tidy(fit2)
```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
# I need more help to fully understand this one. 
#data.mean <- with(fit2, data.frame(rater = yes, location = Toronto, language = English))
#predicted <- cbind(predict(fit2, newdata = data.mean, type = "probs"), data.mean)
#predicted
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


