---
title: "Homework 4: Logistic regression and judges"
author: "your name here"
date: "Due: Sunday, April 19, 2015"
output: html_document
---

This assignment is based on an assignment created by [German Rodriguez](http://data.princeton.edu/wws509/sets/set3.html).  Topics covered include:

- reading in tabular data
- ggplot2
- logistic regression

For this assignment you will be extending the analysis in this article:

Greene and Schaffer (1992) ["Leave to Appeal and Leave to Commence Judicial Review in Canada's Refugee-Determination System: Is the Process Fair?"](http://yorkspace.library.yorku.ca/xmlui/handle/10315/7864) International Journal of Refugee Law, 4.1: 71-83.

Here's how the authors describe the paper:
"The purpose of this study is to determine whether refugees and other persons who are applying for leave to appeal a decision of the Convention Refugee Determination Division of the Canadian Immigration and Refugee Board, a decision of the Appeal Division of the Immigration and Refugee Board, or to commence an action for judicial review have a relatively equal chance to convince the Federal Court of Appeal of the merits of their applications.""

The data and [documentation](http://socserv.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/Greene.pdf) were made avaialble by John Fox.  I have stored a copy of the data in Greene.txt in the subfolder data.


```{r}
library(dplyr)
data <- read.table("data/Greene.txt", header=TRUE)
data <- tbl_df(data)
```

Before doing anything else, just play with the data to get a sense of it.

(@) The main outcome of interest is *decision* which records whether the judge granted an appeal or not.  In what proprotion of cases did the judge grant an appeal.

```{r}
# convert decision to a number rather a factor
data$appeal_granted <- as.numeric(data$decision=="yes")
summarize(data, yes_rate = mean(appeal_granted))
```

(@) There are 12 different judges in the data. A key question is whether different judges have different rates of granting an appeal.  Make a plot showing the number of cases per judge.

```{r}
library(ggplot2)
p <- ggplot(data, aes(judge))
p + geom_bar(stat="bin")
```

(@) Now plot the rate at which the judges grant appeals. 

```{r}
data_to_plot <- group_by(data, judge) %>% 
                  summarize(yes_rate = mean(appeal_granted)) %>%
                  arrange(desc(yes_rate))
data_to_plot$judge <- reorder(data_to_plot$judge, data_to_plot$yes_rate)
p <- ggplot(data_to_plot, aes(x = yes_rate, y = judge))
p + geom_point()
```

(@) Now let's try this with logistic regression because we are going to move to more complicated models.  Fit a logistic regression predicting whetheran appeal will be granted using judge as a categorical predictor. Use Iacobucci as the reference judge to facilitate the comparisons that follow. For more on how to control the reference level of a factor, check out this post on StackOverflow: http://stackoverflow.com/questions/3872070/how-to-force-r-to-use-a-specified-factor-level-as-reference-in-a-regression

```{r}
data$judge <- relevel(data$judge, ref="Iacobucci")
fit <- glm(appeal_granted ~ judge, data = data, family = "binomial")
library(broom)
tidy(fit)
```

(@) Using the results of the model calculate the predicted probability that Judge Marceau will grant an appeal.

```{r}
fit_tidy <- tidy(fit)
predicted_probability <- 1 / (1 + exp(-(fit_tidy[fit_tidy$term=="(Intercept)", "estimate"] + fit_tidy[fit_tidy$term=="judgeMarceau", "estimate"])))
print(paste("Predicted probability for Judge Marceau is:", predicted_probability))
```

(@) Perhaps the different judges have different types of cases.  That might be why they seem to be making such different decisions.  Fit a model treating judge as a factor plus the following predictors: (i) an indicator for whether the expert rater thought the case had merit, (ii) location indicators using 'other' as the reference category, (iii) a language dummy with French as omitted category.  Note that as above you might have to relevel the factors to get this to work as you want.

```{r}
data$location <- relevel(data$location, ref="other")
data$language <- relevel(data$language, ref="French")
fit <- glm(appeal_granted ~ judge + rater + location + language, data = data, family = "binomial")
tidy(fit)
```

(@) For a case that was ruled to have merit, that took place in Toronto, and that was in English, what is the predicted probability that Judge Marceau will grant the appeal?

```{r}
fit_tidy <- tidy(fit)
predictor_value <- fit_tidy[fit_tidy$term=="(Intercept)", "estimate"] +
  fit_tidy[fit_tidy$term=="rateryes", "estimate"] +
  fit_tidy[fit_tidy$term=="judgeMarceau", "estimate"] +
  fit_tidy[fit_tidy$term=="locationToronto", "estimate"] +
  fit_tidy[fit_tidy$term=="languageEnglish", "estimate"] 
predicted_probability <- 1 / (1 + exp(-(predictor_value)))
print(paste("The predicted probability for this case is:", predicted_probability))
```


#### The command below is helpful for debugging, please don't change it

```{r echo=FALSE}
sessionInfo()
```


